A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Kamala Harris Emerges as a Leading Candidate with Constant Support Rates Hovering Around 50%

### 2024-10-25 paper/paper.qmd
update 'correlation between predictor variables'

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(broom)
library(modelsummary)
library(rstanarm)
library(splines)
library(ggplot2)
library(knitr)
library(kableExtra)
library(gridExtra)
library(here)

@@

## Outcome variables

### pct

@@

### end_date

@@

The date variable representing the time the poll was concluded helps was put into account to keep track of how support for a candidate changes in time. The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 1 January, 2023 to 9 September, 2024 (@tbl-date). @fig-date shows that the polling data is more concentrated on survey results conducted in the recent period.

@@

### pollster, state

@@

### pollscore

@@

>>>>

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(broom)
library(modelsummary)
library(rstanarm)
library(splines)
library(ggplot2)
library(knitr)
library(kableExtra)
library(gridExtra)
library(here)
library(usmap)  # For the map of US states
library(viridis)  # For a color palette

@@

## Outcome variable

### The Proportion(%) of Support that a Candidate Received in the Poll

@@

### The Date the Poll was Concluded

@@

The 'end_date' variable representing the time the poll was concluded was put into account to keep track of how support for a candidate changes in time. The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 1 January, 2023 to 9 September, 2024 (@tbl-date). @fig-date shows that the polling data is more concentrated on survey results conducted in the recent period.

@@

### Pollster and State

@@

### Pollscore Measuring the Validity of Polling Questions

@@

## Correlation Between Predictor Variables

By examining the relationships between predictor variables in the analysis dataset, we aim to identify potential biases in interpretation and model instability. This is expected to ensure a transparent understanding of the model's robustness and interpretability.

### End Date and State

@fig-correlation1 shows notable polling concentrations on certain states such as Pennsylvania, Ohio, and Nevada in week 2024-37 (September 9, 2024) or 2024-38 (September 15, 2024). This suggests that these states are key battelgrounds or areas of strategic focus during the election period. Moreover, polling activities are not consistent across all weeks and there are clear peaks in polling activity in week 2024-38 (September 15, 2024), which may correspond to significant political events, debates, or media focuses.

```{r}
#| label: fig-correlation1
#| fig-cap: "Polling concentration over time by state *Note:* Strategic focus of polling is directed in battleground states such as Pennsylvania and Ohio in September 15, 2024"
#| echo: false

# Aggregate the number of polls by state and by week (or month)
poll_time_state_2 <- harris_data %>%
  mutate(week = format(as.Date(end_date), "%Y-%U")) %>%  # Group by week of the year
  group_by(state, week) %>%
  summarise(total_polls = n(), .groups = "drop")  # Adding .groups = "drop"

# Plot a heatmap of polling activity by state and week
ggplot(poll_time_state_2, aes(x = week, y = state, fill = total_polls)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
  labs(x = "Week",
       y = "State",
       fill = "Total Polls") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

### End Date and Pollster

@fig-correlation2 shows that Siena/NYT, which made up the largest proportion of polls in our analysis dataset, has concentrated polling efforts in week 2024-37 (September 9, 2024) and week 2024-38 (September 15, 2024). Notably active periods, possibly around key election events, might introduce temporal bias and overemphasize methodologies of Siena/NYT in that specific time period. Emerson shows overall consistency of polling activities across several weeks indicating steady involvement, while other polling organizations show more sporadic or minimal involvement.

```{r}
#| label: fig-correlation2
#| fig-cap: "Polling concentration over time by pollster *Note:* Concentrated polling efforts of Siena/NYT in the first half of September, 2024 show temporal bias and potential overemphasizing of their polling methodologies around key election events."
#| echo: false

# Aggregate the number of polls by state and by week (or month)
poll_time_pollster <- harris_data %>%
  mutate(week = format(as.Date(end_date), "%Y-%U")) %>%  # Group by week of the year
  group_by(pollster, week) %>%
  summarise(total_polls = n(), .groups = "drop")  # Adding .groups = "drop"

# Plot a heatmap of polling activity by state and week
ggplot(poll_time_pollster, aes(x = week, y = pollster, fill = total_polls)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
  labs(x = "Week",
       y = "Pollster",
       fill = "Total Polls") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

### Pollster and State

@fig-correlation3 shows that Siena/NYT has the strongest state-level presence in key battleground states like Michigan and Arizona. Others like AtlasIntel and Beacon/Shaw has minimal polling coverage in targeted states such as Nevada and Georgia. This vairation in coverage could influence the overall analysis in that some states might receive disproportionately more attention from certain pollsters.

```{r}
#| label: fig-correlation3
#| fig-cap: "Polling activity by pollster and state *Note:* Siena/NYT has strongest state-level polling coverage in key battleground states"
#| echo: false

# Step 1: Filter out NAs from pollster column and summarize data by state and pollster
pollster_state_data <- harris_data %>%
  filter(!is.na(pollster)) %>%  # Remove NAs from pollster
  group_by(state, pollster) %>%
  summarise(total_polls = n(), .groups = "drop")  # Summarize number of polls

# Step 2: Plot the faceted map using plot_usmap
plot_usmap(data = pollster_state_data, values = "total_polls", regions = "states") +
  facet_wrap(~ pollster, ncol = 2) +  # Create a map for each pollster
  scale_fill_distiller(palette = "Blues", direction = 1, na.value = "gray90") +  # Use blue color palette
  labs(fill = "Total Polls") +
  theme_void() +  # Remove axes and gridlines
  theme(strip.text = element_text(size = 12))  # Customize facet labels
